---
title: "Methods and Initial Results"
author: "Nora Nickels"
date: "5/9/2018"
header-includes:
    - \usepackage{setspace}
    - \usepackage{booktabs}
    - \usepackage{longtable}
    - \usepackage{array}
    - \usepackage{multirow}
    - \usepackage[table]{xcolor}
    - \usepackage{wrapfig}
    - \usepackage{float}
    - \usepackage{colortbl}
    - \usepackage{pdflscape}
    - \usepackage{tabu}
    - \usepackage{threeparttable}
    - \usepackage{threeparttablex}
    - \usepackage[normalem]{ulem}
    - \usepackage{makecell}
output: pdf_document
---
\doublespacing
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(knitr.table.format = "latex")

library(tidyverse)
library(ggplot2)
library(knitr) 
library(modelr)
library(broom)
library(kableExtra)

set.seed(1234)
theme_set(theme_minimal())

macss <- read_csv('macss.csv')
desc1 <- read_csv('macss_desc_1.csv')
desc2 <- read_csv('macss_desc_2.csv')
```

# Methods

## Theoretical Model
* Define the model (but don't reproduce the literature review - you'll link these together in the final paper)

### Variables
* Map variables to theoretical constructs

## Data

### Data collection
* Describe the data and how to obtain it
* Describe data collection procedures

### Participants

### Summary Statistics
* Summary statistics for key variables

```{r desc_1}

kable(desc1, format = "latex", caption = "Demographics [note]", booktabs = T) %>%
  kable_styling(latex_options = "striped") %>%
  column_spec(5, width = "5em") %>%
  column_spec(6, width = "5em") %>%
  column_spec(7, width = "5em") %>%
  add_footnote(c("Caption Notes"))

```

```{r desc_2}

kable(desc2, format = "latex", 
      caption = "Descriptive Statistics of TAI and PANAS Inventories [note]",
      booktabs = T) %>%
  kable_styling(latex_options = "striped") %>%
  column_spec(5, width = "5em") %>%
  column_spec(6, width = "5em") %>%
  column_spec(7, width = "5em") %>%
  add_footnote(c("Caption Notes"))

```


### Data Analysis
* Define the methods used to analyze the data

# Results
* Present initial (tentative results)
* Use tables and figures as appropriate
