---
title: "Problem Set 2: Interaction"
author: "Nora Nickels"
date: "4/26/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Problem Set 2

## Problem 2: Interactions Terms

```{r data}
# Load libraries

library(tidyverse)
library(forcats)
library(broom)
library(modelr)
library(stringr)
library(titanic)
library(rcfss)
library(car)
library(haven)

options(digits = 3)
set.seed(1234)
theme_set(theme_minimal())


# Load data

biden_df = read.csv("biden.csv") %>% 
  na.omit() %>%
  mutate(dem = factor(dem),
         rep = factor(rep))

```

### Linear Regression Model

```{r LRM}

# Fit linear regression model of biden score on age, education and the interaction b/t age and educ.
lm_biden <- biden_df %>%
  lm(biden ~ age + educ + age * educ, data = .)

# Report regression coefficients  
tidy(lm_biden)

biden_coef <- tidy(lm_biden)

# Report goodness of fit
glance(lm_biden)

biden_fit <- glance(lm_biden)

```

### 2A: Evaluate the marginal effect of age on Joe Biden thermometer rating, conditional on education

```{r age}

# function to get point estimates and standard errors
# model - lm object
# mod_var - name of moderating variable in the interaction
instant_effect <- function(model, mod_var){
  # get interaction term name
  int.name <- names(model$coefficients)[[which(str_detect(names(model$coefficients), ":"))]]
  
  marg_var <- str_split(int.name, ":")[[1]][[which(str_split(int.name, ":")[[1]] != mod_var)]]
  
  # store coefficients and covariance matrix
  beta.hat <- coef(model)
  cov <- vcov(model)
  
  # possible set of values for mod_var
  if(class(model)[[1]] == "lm"){
    z <- seq(min(model$model[[mod_var]]), max(model$model[[mod_var]]))
  } else {
    z <- seq(min(model$data[[mod_var]]), max(model$data[[mod_var]]))
  }
  
  # calculate instantaneous effect
  dy.dx <- beta.hat[[marg_var]] + beta.hat[[int.name]] * z
  
  # calculate standard errors for instantaeous effect
  se.dy.dx <- sqrt(cov[marg_var, marg_var] +
                     z^2 * cov[int.name, int.name] +
                     2 * z * cov[marg_var, int.name])
  
  # combine into data frame
  data_frame(z = z,
             dy.dx = dy.dx,
             se = se.dy.dx)
}

# Plot point range plot conditional on education

instant_effect(lm_biden, "educ") %>%
  ggplot(aes(z, dy.dx,
             ymin = dy.dx - 1.96 * se,
             ymax = dy.dx + 1.96 * se)) +
  geom_pointrange() +
  geom_hline(yintercept = 0, linetype = 2) +
  labs(title = "Marginal effect of age",
       subtitle = "By education",
       x = "Education",
       y = "Estimated marginal effect")

# Plot line plot conditional on education

instant_effect(lm_biden, "educ") %>%
  ggplot(aes(z, dy.dx)) +
  geom_line() +
  geom_line(aes(y = dy.dx - 1.96 * se), linetype = 2) +
  geom_line(aes(y = dy.dx + 1.96 * se), linetype = 2) +
  geom_hline(yintercept = 0) +
  labs(title = "Marginal effect of age",
       subtitle = "By education",
       x = "Education",
       y = "Estimated marginal effect")

# Test statistical significance conditional on education

linearHypothesis(lm_biden, "age + age:educ")

```

Based on the above plots and linear hypothesis testing, we can conclude that there is a significant marginal effect of age on biden rating, conditional on education. The marginal effects is positive for individuals over 14 years of education and negative for individuals under 14 years of education For individuals who have 13 - 16 years of education roughly, we can't confidently say that the marginal effect differs from zero.


### 2B: Evaluate the marginal effect of education on Joe Biden thermometer rating, conditional on age

```{r education}
# marginal effect

# Plot point range plot conditional on age

instant_effect(lm_biden, "age") %>%
  ggplot(aes(z, dy.dx,
             ymin = dy.dx - 1.96 * se,
             ymax = dy.dx + 1.96 * se)) +
  geom_pointrange() +
  geom_hline(yintercept = 0, linetype = 2) +
  labs(title = "Marginal effect of education",
       subtitle = "By age",
       x = "Age",
       y = "Estimated marginal effect")

# Plot line plot conditional on age

instant_effect(lm_biden, "age") %>%
  ggplot(aes(z, dy.dx)) +
  geom_line() +
  geom_line(aes(y = dy.dx - 1.96 * se), linetype = 2) +
  geom_line(aes(y = dy.dx + 1.96 * se), linetype = 2) +
  geom_hline(yintercept = 0) +
  labs(title = "Marginal effect of education",
       subtitle = "By age",
       x = "Age",
       y = "Estimated marginal effect")

# Test statistical significance conditional on education

linearHypothesis(lm_biden, "educ + age:educ")


```

Based on the above plots and linear hypothesis testing, we can conclude that there is a significant marginal effect of education on biden rating, conditional on age. However, the marginal effect is only significant for individuals who are over 45 years of age; for these individuals, the marginal effect of education on biden rating is negative. 

